---
title: 'Assignment 2: Data visualization'
author: "Tamas Nagy"
output: html_document
editor_options: 
  chunk_output_type: console
---

You will have to create 3 plots based on the datasets and instructions detailed below. You will find the plots themeselves in the `assignments/assignment_2_plots`. Your task is to write the code that will reproduce the plots as closely as possible.

# Skills needed to solve this assignment

-   Using R and RStudio, reading data
-   Reporting using RMarkdown
-   Using Git and Github (for submitting the task)
-   Data manipulation (e.g. dplyr, tidyr), and working with factors (forcats)
-   Data visuzlization (ggplot2)

```{r setup}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(tidytuesdayR)
library(readr)
library(viridis)
library(scales)
library(RColorBrewer)
```

## Task 1: Climbing expeditions

The 2020-09-22 TidyTueday datasets are about climbing expeditions. From the three datasets, use the "expeditions". Reproduce the plot below! Notice a few things:

-   Use `forcats::fct_lump()` to get the 15 most frequent peaks, and drop the "Other" category.
-   The bars are ordered by the sum of all expeditions (use `fct_reorder()`).
-   The bar colors use the viridis palette and light theme.

```{r}
expeditions <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-22/expeditions.csv') %>% 
  mutate(peak_name = fct_lump(peak_name, n = 15)) %>%
  filter(peak_name != "Other") %>% 
  group_by(peak_name, season) %>%
  summarize(count = n())
  

ggplot(expeditions, aes(x = count, y = reorder(peak_name, count), fill = season)) +
  geom_bar(stat = "identity") +
  labs(title = "Expeditions by Season and Peak",
       x = "Number of Expeditions",
       y = "Peak Name") +
  scale_colour_discrete(palette = viridis) +
  theme_light()

```

## Task 2: PhDs awarded

The 2019-02-19 TidyTueday dataset is about phd-s awarded by year and field. There is only one dataset, it is called `phd_by_field`. Reproduce the plot below!

Notes:

-   First you have to aggregate the data to count all phd-s by broad fields.
-   To make the x axis breaks pretty, use `scales::pretty_breaks()`, to make the y axis labels comma formatted, use `scales::comma_format()`.
-   The line size is 1.2, the colors are from the brewer "Dark2" palette. The theme is set to minimal.

```{r}
phd_field <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-02-19/phd_by_field.csv") %>%
  na.omit() %>% 
  group_by(broad_field, year) %>% 
  summarize(n = sum(n_phds))

ggplot(phd_field, aes(x = year, y = n, group = broad_field, color = broad_field)) +
  geom_line(linewidth = 1.2) +
  scale_color_brewer(palette = "Dark2")+
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_y_continuous(labels = comma_format()) +
  theme_minimal()
  
```

## Task 3: Commute in the US

The 2019-11-05 TidyTueday dataset is about commuting to work in each city in the US by bike or on foot. There is only one dataset, it is called `commute`. Reproduce the plot below!

Notes:

-   First you have to aggregate the data to count all commutes by state.
-   Both axis scales are log transformed and the labels comma formatted, using `scales::comma_format()`
-   The point size is 2, . The theme is set to light.

```{r}
commute_mode <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-11-05/commute.csv") %>% 
  na.omit() %>% 
  group_by(state, state_abb, state_region, mode) %>% 
  summarize(n = sum(n)) %>% 
  pivot_wider(names_from = mode, values_from = n)

ggplot(commute_mode, aes(x = log(Walk), y = log(Bike), color = state_region)) +
  geom_point(size = 2) +
  geom_text(aes(label = state_abb), size = 3, nudge_x = 0.1, nudge_y = 0.1) +
  scale_x_continuous(labels = comma_format()) +
  scale_y_continuous(labels = comma_format()) +
  theme_light()
```
